---
title: "World Happiness"
author: "Amy Mailo"
date: "2022-12-01"
output:
  pdf_document: default
  html_document: default
---

1. Import the data 
```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(forcats)
library(lubridate) 
library(ggplot2)
```

```{r}
WH <-read_csv("WorldHappiness_Corruption_2015_2020.csv")
#https://www.kaggle.com/datasets/eliasturk/world-happiness-based-on-cpi-20152020
```

2. Present the data as a data frame and a tibble. 
```{r}
#Set WH as a data frame and as a tibble. 
WH_data_frame <-as.data.frame(WH)
WH_tibble <-as_tibble(WH) 
is.tibble(WH_tibble) 
```

Use head() function to see which columns are doubles and which are characters. For this data set, all the character columns (Country and continent) are categorical variables and all the double columns are quantitative (the rest of the columns). 
```{r}
#take a glimpse at the data set 
head(WH)
dim(WH)
summary(WH)
```
The data set WH has 792 rows and 13 columns about the details of each country's happiness score. The summary() function gives us more details about each of the variables.

3. Explain the data 

The WH data file contains data from 2015-2020 on Happiness and Corruption data. More specifically this data comes from two sources: 1) The World Happiness Report (data gathered from the Gallup World Poll surveys) 2) Corruption Perceptions Index (data gathered from Transparency International). 

Define the names and columns. 
1. **Country**: This is a categorical variable that describes the name of country. 
2. **happiness_score**: This is a quantitative variable that describes the average of country's happiness score to the primary life evaluation question from the Gallup World Poll.
3. **gdp_per_capital**: This is a quantitative variable that describes the extent to which GDP contributes to the calculation of the Happiness Score.
4. **family**: This is a quantitative variable that describes the extent to which Family contributes to the calculation of the Happiness Score.
5. **health**: This is a quantitative variable that describes the extent to which Life expectancy contributes to the calculation of the Happiness Score.
6. **freedom**: This is a quantitative variable that describes the extent to which Freedom contributes to the calculation of the Happiness Score.
7. **generosity**: This is a quantitative variable that describes a numerical value calculated based on poll participants' perceptions of generosity in their country.
8. **government_trust**: This is a quantitative variable that describes the extent to which Perception of Corruption contributes to Happiness Score.
9. **dystopia_residual**: This is a quantitative variable that describes a score based on a hypothetical comparison to the world's saddest country.
10. **continent**: This is a categorical variable that describes the region of the country. 
11. **Year**: This is a quantitative variable that describes the date for each of the findings.
12. **social_support**: This is a quantitative variable that describes people of each country's perception and actuality that one is cared for, has assistance available from other people, and that one is part of a supportive social network.
13. **cpi_score**: This is a quantitative variable that describes the corruption perception index. The higher the score, the better (meaning the country's respondents perceive less corruption).


4. Tidy the data
Use tidy R and dplyr functions to modify the data so that it is structured properly for better analysis and processing.

First, I rename the columns to shorten the names and standardize capitalization.
```{r}
WH%>%
  rename("Happiness" = "happiness_score")%>%
  rename("GDPperCapita" = "gdp_per_capita")%>%
  rename("Family" = "family")%>%
  rename("Health" = "health")%>%
  rename("Freedom" = "freedom")%>%
  rename("Generosity" = "generosity")%>%
  rename("TrustinGovernment" = "government_trust")%>%
  rename("Dystopia" = "dystopia_residual")%>%
  rename("Continent" = "continent")%>%
  rename("SocialSupport" = "social_support")%>%
  rename("CPI" = "cpi_score")->
WH1
WH1 
```

Next, I take a closer look at the levels of the categorical variables to see if further tidying was needed.
```{r}
library(forcats)
levels(factor(WH1$Country))
levels(factor(WH1$Continent))
```
This factoring demonstrated that there were several countries in the dataset. However, the factors that I would use to tidy some of the entries is already captured by the Continent column variable, which identifies which continent each country belongs to. So, when it comes to analysis, I will focus on the country and the continents. 
 
Then, I list the unique observations of each of the categorical columns to get a better sense of the data I am looking at. 
```{r}
WH1%>%
  count(Year) %>%
  filter(n>1)%>%
  arrange(desc(n))
#132 observations for each of the years 2015-2020

WH1%>%
  count(Country) %>%
  filter(n>1)%>% 
  arrange(desc(n))
#All countries listed 6 times each 

WH1%>%
  count(Continent) %>%
  filter(n>1)%>%
  arrange(desc(n)) 
#Observations of continent they belong to (in order of frequency): 1) Europe 2) Asia 3) Africa 4) South America 5) North America 6) Australia  
```

The observations for Country, Continent, and Year are NOT keys since their observations repeat. To more easily analyze and visualize the data, I use group_by() and summarize() to take the average of the column variable scores from 2015 to 2020 for each of the countries/continents.
```{r}
WH1 %>%
  select(Continent , Happiness, GDPperCapita, Family, Health, Freedom, Generosity, TrustinGovernment, Dystopia, SocialSupport, CPI)%>%
  group_by(Continent) %>%
  summarise(MeanHappiness = mean(Happiness), 
            MeanGDPperCapita = mean(GDPperCapita), 
            MeanFamily = mean(Family), 
            MeanHealth = mean(Health), 
            MeanFreedom = mean(Freedom), 
            MeanGenerosity = mean(Generosity), 
            MeanTrustinGovernment = mean(TrustinGovernment), 
            MeanDystopia = mean(Dystopia), 
            MeanSocialSupport = mean(SocialSupport), 
            MeanCPI = mean(CPI))-> WH1Continent
WH1Continent

WH1 %>%
  select(Country , Happiness, GDPperCapita, Family, Health, Freedom, Generosity, TrustinGovernment, Dystopia, SocialSupport, CPI)%>%
  group_by(Country) %>%
  summarise(MeanHappiness = mean(Happiness), 
            MeanGDPperCapita = mean(GDPperCapita), 
            MeanFamily = mean(Family), 
            MeanHealth = mean(Health), 
            MeanFreedom = mean(Freedom), 
            MeanGenerosity = mean(Generosity), 
            MeanTrustinGovernment = mean(TrustinGovernment), 
            MeanDystopia = mean(Dystopia), 
            MeanSocialSupport = mean(SocialSupport), 
            MeanCPI = mean(CPI))-> WH1Country
WH1Country
```


Use ggplot coding to generate plots that convey interesting and impact facts and trends of selected categorical and quantitative variables.

First, I start with data visualizations to explore the Country and Continent variable. Initially, I wanted to look at the U.S. to narrow down the observations, but other countries or regions may also yield interesting results. For the second part of my data visualization, I will focus on the variables in the U.S. and/or certain continents. 
```{r}
ggplot(data = WH1Country) +
  geom_point(mapping = aes(x = Country, y = MeanHappiness))

ggplot(data = WH1Continent) +
  geom_point(mapping = aes(x = Continent, y = MeanHappiness))
#The continents Australia and North American seem to have a higher happiness score with less variability to their responses. 
#maybe a bar graph would be a better representation in this case? 
ggplot(data = WH1Continent) +
  geom_bar(mapping = aes(x = Continent, y=MeanHappiness, fill = Continent), stat = "identity")


#Who are the 5 MOST happy countries? 
WH1Country%>%
  arrange(desc(MeanHappiness))-> W10MostHappy
  head(W10MostHappy, 5)-> MostHappy
MostHappy

  ggplot(data = MostHappy) +
  geom_bar(mapping = aes(x = Country, y=MeanHappiness, fill = Country), stat = "identity")
#Which continents are the MOST happy? 
  WH1Continent%>%
  arrange(desc(MeanHappiness))-> ContinentMostHappy
ContinentMostHappy

  ggplot(data = ContinentMostHappy) +
  geom_bar(mapping = aes(x = Continent, y=MeanHappiness, fill = Continent), stat = "identity")

#Which 5 countries have the LEAST amount of corruption? 
WH1Country%>%
  arrange(desc(MeanCPI))-> W10LeastCorrupt
  head(W10LeastCorrupt, 5)-> LeastCorrupt
LeastCorrupt

ggplot(data = LeastCorrupt) +
  geom_bar(mapping = aes(x = Country, y=MeanCPI, fill = Country), stat = "identity")
#Which continents have the LEAST amount of corruption? 
WH1Continent%>%
  arrange(desc(MeanCPI))-> ContinentLeastCorrupt
ContinentLeastCorrupt

ggplot(data = ContinentLeastCorrupt) +
  geom_bar(mapping = aes(x = Continent, y=MeanCPI, fill = Continent), stat = "identity")


#Who are 5 LEAST happy countries? 
WH1Country%>%
  arrange(MeanHappiness)-> W10LeastHappy
  head(W10LeastHappy, 5)-> LeastHappy
LeastHappy

  ggplot(data = LeastHappy) +
  geom_bar(mapping = aes(x = Country, y=MeanHappiness, fill = Country), stat = "identity")
#Who are least happy continents?
WH1Continent%>%
  arrange(MeanHappiness)-> ContinentLeastHappy
ContinentLeastHappy

  ggplot(data = ContinentLeastHappy) +
  geom_bar(mapping = aes(x = Continent, y=MeanHappiness, fill = Continent), stat = "identity")
  
#Which 5 countries have the MOST amount of corruption?
WH1Country%>%
  arrange(MeanCPI)-> W10MostCorrupt
  head(W10MostCorrupt, 5)-> MostCorrupt
MostCorrupt

ggplot(data = MostCorrupt) +
  geom_bar(mapping = aes(x = Country, y=MeanCPI, fill = Country), stat = "identity")

#Which continents have the MOST amount of corruption?
WH1Continent%>%
  arrange(MeanCPI)-> ContinentMostCorrupt
ContinentMostCorrupt

ggplot(data = ContinentMostCorrupt) +
  geom_bar(mapping = aes(x = Continent, y=MeanCPI, fill = Continent), stat = "identity") 
```

My main question is which predictor variable has the strongest association with Happiness. Accordingly, I look at the data visualizations between the following variables and make the following guesses: 
- **Happiness** and **GDPperCapita** where as GDP per Capita increases, Happiness increases. 
- **Happiness** and **Family** where as Family score increases, Happiness increases. 
- **Happiness** and **Health** where as Health increases, Happiness increases. 
- **Happiness** and **Freedom** where as Freedom increases, Happiness increases. 
- **Happiness** and **Generosity** where as Generosity increases, Happiness increases 
- **Happiness** and **GovernmentTrust** where as Government Trust increases, Happiness increases 
- **Happiness** and **Dystopia** where as Dystopia increases, Happiness increases
- **Happiness** and **SocialSupport** where as the SocialSupport increases, Happiness increases 
- **Happiness** and **CPI** where as the CPI score increases (meaning less corruption), Happiness increases 

```{r}
#MeanGDPperCapita  
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanGDPperCapita, y = MeanHappiness), stat = "identity")+
  xlab("Mean GDP per Capita")+
  ylab("Mean Happiness")+
  ggtitle("Mean GDP per Capita and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanGDPperCapita, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanGDPperCapita,WH1Country$MeanHappiness)
#Returns 0.82 which indicates a fairly positive relationship 
lm(Happiness~GDPperCapita, data=WH1)->lmGDP
lmGDP
summary(lmGDP) 
# happiness score = 3.32 + GDPperCapita(2.31)
#On average, for every one point increase in GDPperCapita score, the happiness increases by 3.21 (scale of 0-10). 

#MeanFamily 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanFamily, y = MeanHappiness), stat = "identity")+
  xlab("Mean Family")+
  ylab("Mean Happiness")+
  ggtitle("Mean Family and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanFamily, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanFamily,WH1Country$MeanHappiness)
#Returns 0.75 which indicates a moderately positive relationship 
lm(Happiness~Family, data=WH1)->lmFamily
lmFamily
summary(lmFamily)
# happiness score = 5.31 + Family(0.32)
#On average, for every one point increase in Family score, the happiness increases by 0.32 (scale of 0-10). 


#MeanHealth 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanHealth, y = MeanHappiness), stat = "identity")+
  xlab("Mean Health")+
  ylab("Mean Happiness")+
  ggtitle("Mean Health and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanHealth, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanHealth,WH1Country$MeanHappiness)
#Returns 0.80 which indicates a fairly positive relationship 
lm(Happiness~Health, data=WH1)->lmHealth
lmHealth
summary(lmHealth)
# happiness score = 3.11 + Health(3.64)
#On average, for every one point increase in Health score, the happiness increases by 4.64 (scale of 0-10). 


#MeanFreedom 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanFreedom, y = MeanHappiness), stat = "identity")+
  xlab("Mean Freedom")+
  ylab("Mean Happiness")+
  ggtitle("Mean Freedom and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanFreedom, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanFreedom,WH1Country$MeanHappiness)
#Returns 0.58 which indicates a moderately positive relationship 
lm(Happiness~Freedom, data=WH1)->lmFreedom
lmFreedom
summary(lmFreedom)
# happiness score = 3.71 + Freedom(4.14)
#On average, for every one point increase in Freedom score, the happiness increases by 4.14 (scale of 0-10). 


#MeanGenerosity 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanGenerosity, y = MeanHappiness), stat = "identity")+
  xlab("Mean Generosity")+
  ylab("Mean Happiness")+
  ggtitle("Mean Generosity and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanGenerosity, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanGenerosity,WH1Country$MeanHappiness)
#Returns 0.17 which indicates a negligible positive relationship 


#MeanTrustinGovernment 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanTrustinGovernment, y = MeanHappiness), stat = "identity")+
  xlab("Mean Government Trust")+
  ylab("Mean Happiness")+
  ggtitle("Mean Government Trust and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanTrustinGovernment, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanTrustinGovernment,WH1Country$MeanHappiness)
#Returns 0.47 which indicates a weak positive relationship 


#MeanDystopia 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanDystopia, y = MeanHappiness), stat = "identity")+
  xlab("Mean Dystopia")+
  ylab("Mean Happiness")+
  ggtitle("Mean Dystopia and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanDystopia, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanDystopia,WH1Country$MeanHappiness)
#Returns 0.53 which indicates a weak positive relationship 


#MeanSocialSupport 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanSocialSupport, y = MeanHappiness), stat = "identity")+
  xlab("Mean Social Support")+
  ylab("Mean Happiness")+
  ggtitle("Mean Social Support and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanSocialSupport, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanSocialSupport,WH1Country$MeanHappiness)
#Returns 0.79 which indicates a moderately positive relationship 
lm(Happiness~SocialSupport, data=WH1)->lmSS
lmSS
summary(lmSS)
# happiness score = 5.26 + Social Support(0.34)
#On average, for every one point increase in Social Support score, the happiness increases by 0.34 (scale of 0-10). 
#Social Support had a lot of 0s which may be of concern since the dataset automatically filled in missing values as 0. Due to the overwhelming amount of 0s for Social Support, I will err on the side of caution


#MeanCPI 
ggplot(data = WH1Country)+
  geom_point(mapping = aes(x = MeanCPI, y = MeanHappiness), stat = "identity")+
  xlab("Mean CPI")+
  ylab("Mean Happiness")+
  ggtitle("Mean CPI and Mean Happiness 2015-2020")+
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.5))+
  geom_smooth(aes(x = MeanCPI, y = MeanHappiness),method = "lm", col = "dark green", se=FALSE)
cor(WH1Country$MeanCPI,WH1Country$MeanHappiness)
#Returns 0.71 which indicates a moderately positive relationship 
lm(Happiness~CPI, data=WH1)->lmCPI
lmCPI
summary(lmCPI)
# happiness score = 3.70 + CPI(0.04)
#On average, for every one point increase in CPI score (scale of 0-100), the happiness increases by 0.04 (scale of 0-10). 
```


#Data Analysis Considerations and Notes
*If you add up the factors (GDP per Capita, Family, Life Expectancy, Freedom, Generosity, Trust in Government Corruption), then you will get the happiness score – so it may be unreliable to model these variables to predict Happiness scores 
- That is why we looked at Happiness and CPI -- and why that relationship should be highlighted 

*The dataset had already plugged in 0’s for the missing values 
